A Comparative study of Dask and Spark for neuroimaging pipelines

### 2019-06-26 latex/paper.tex
Worked on the engine section

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage{times}
\usepackage{siunitx}
\usepackage{latexsym}
\usepackage{hyperref}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{caption}
\usepackage{subcaption}
\algnewcommand\algorithmicforeach{\textbf{for each}}
\algdef{S}[FOR]{ForEach}[1]{\algorithmicforeach\ #1\ \algorithmicdo}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

@@

\title{Comparative study of Dask and Spark for neuroimaging pipelines\\
{\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and
should not be used}
\thanks{Natural Sciences and Engineering Research Council}
}

\author{\IEEEauthorblockN{1\textsuperscript{st} Mathieu Dugr\'e}
\IEEEauthorblockA{\textit{Computer Science and Software Engineering} \\
\textit{Concordia University}\\
Montreal, Canada \\
mathieu.dugre@mail.concordia.ca}
\and
\IEEEauthorblockN{2\textsuperscript{nd} Given Name Surname}
\IEEEauthorblockA{\textit{Computer Science and Software Engineering} \\
\textit{Concordia University}\\
Montreal, Canada \\
email address}
\and
\IEEEauthorblockN{3\textsuperscript{rd} Given Name Surname}
\IEEEauthorblockA{\textit{Computer Science and Software Engineering} \\
\textit{Concordia University}\\
Montreal, Canada \\
email address}
}

@@

%Similarities
Spark and Dask offer in-memory computing, data locality, and lazy evaluation; which
is common for Big Data framework. Both their scheduler operates dynamically. This is
good when the runtimes are not known ahead of time~\cite{Dask:15}. Over these
similarities, the frameworks are quite different.

@@

\subsection{Engines}
For all our experiments, we used Spark v2.4.0 and Dask v1.1.4.

\subsubsection{PySpark RDD~\cite{spark-rdd}} Spark is a general-purpose Big Data
framework. It provides in-memory computing, data locality, and lazy evaluation. Spark
abstracts resilient distributed dataset (RDD)~\cite{RDD} making it fault-tolerant.
Its high-level API provides optimization for transformations applied on data. Since
Spark is written in Scala, this optimization is lessened because it requires a
serialization from Python to Java. The API offered by Spark is flexible and suits
most problems however it differs from the Python ecosystem. The standalone scheduler
is used for our experiment. It has a LIFO policy.

\subsubsection{Dask Bag~\cite{dask-bag}} Dask Bag is one of the high-level API
offered by Dask. Like Spark, it provides in-memory computing, data locality, and lazy
evaluation however it does not have fault resilience. Fundamentally, it is
parallelized lists. Dask Bag parallelizes computations across a large collection of
generic python objects~\cite{dask-bag}. Also, it offers optimized algorithms that
mimic most of the toolz~\cite{toolz} library. Our experiment uses the Dask
Distributed~\cite{dask-distributed} scheduler. It has a LIFO policy. In theory, it
should speed-up the pipeline execution by finishing a branch of task computations
before starting new ones.

\subsubsection{Dask Delayed~\cite{dask-delayed}} Dask Delayed is a low-level API. It
does not offer the optimized algorithm as the high-level APIs however it allows the
user to parallelize custom function that does not fit in the array, dataframe or
MapReduce framework. This is perfect for the more complex functions which could
potentially be parallelized. Like Dask Bag, it offers in-memory computing, data
locality and lacks fault resilience. The Dask Distributed scheduler is also used.

\subsubsection{Dask Futures~\cite{dask-futures}} Dask Futures is also a low-level API
from Dask. Like the other Dask APIs, it benefits of in-memory computing and data
locality. However, unlike the other APIs, Dask Futures triggers computation
immediately. When the user tries to gather data if available it will return the
result otherwise it will block until the computation is done. Again, the Dask
Distributed scheduler is used.

@@

We would like to thanks Compute Canada Cloud for the infrastructure.

>>>>

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage{times}
\usepackage{siunitx}
\usepackage{latexsym}
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor=black,citecolor=blue,filecolor=black,urlcolor=blue}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{caption}
\usepackage{subcaption}

\newcommand{\TG}[1]{\color{blue}From Tristan: #1 \color{black}}

\algnewcommand\algorithmicforeach{\textbf{for each}}
\algdef{S}[FOR]{ForEach}[1]{\algorithmicforeach\ #1\ \algorithmicdo}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

@@

\title{A Comparative study of Dask and Spark for neuroimaging pipelines}

\author{Mathieu Dugr\'e, Val\'erie Hayot-Sasson, Tristan Glatard\\
Department of Computer Science and Software Engineering\\
Montr\'eal, Qu\'ebec, Canada
}

@@

%Similarities
\TG{Note for later: explain the following concepts in the intro: lazy eval, data locality, in-memory computing.}
Spark and Dask offer in-memory computing, data locality, and lazy evaluation; which
is common for Big Data framework. Both their scheduler operates dynamically. This is
good when the runtimes are not known ahead of time~\cite{Dask:15}. Over these
similarities, the frameworks are quite different.

@@

\subsection{Engines}

\subsubsection{Apache Spark} Apache Spark is a general-purpose Big Data
framework. It provides in-memory computing, data locality, and lazy
evaluation. Spark's main data structure, the Resilient Distributed Dataset
(RDD)~\cite{RDD}, is a fault-tolerant and parallel collection of data
elements. 
Dataframes, Spark's other popular API, are an overlay of RDDs with similar 
performance features. 
Spark's core is written in Scala, but it also has an R and a Python API
called PySpark. We used PySpark, given the popularity of Python in
neuroscience, although the serialization cost of Python objects to Scala
might be a substantial source of overhead. We used Spark's default
scheduler, Spark standalone, although YARN~\cite{yarn} and
Mesos~\cite{mesos} are also available. In the Spark standalone scheduler, a
process called the \emph{master} coordinates resources provisioned by
\emph{workers} on the cluster. The application is submitted to the
\emph{driver} that in turn requests workers to the master and dispatches
tasks to them. The Spark standalone scheduler uses a LIFO
(Last-In-First-Out) task scheduling policy \TG{we should describe it
briefly and refer to the paper}. The Spark standalone scheduler has two
execution modes: client mode, where the driver runs in a dedicated  process
on the cluster, and cluster mode, where the driver runs in a worker. We
used client mode as PySpark is not available in cluster mode. We used
Apache Spark v2.4.0.

\subsubsection{Dask} Dask is a Big Data engine that is becoming
increasingly popular in the scientific Python ecosystem. \TG{A few more
introductory sentences should be added, following Spark's model.}
Dask represents application tasks as Dask graphs, produced by a
user-facing API and executed by a scheduler. We used the \href{https://distributed.dask.org/en/latest/index.html}{Dask
Distributed} scheduler.
Dask Distributed also has a LIFO policy. In theory, it should speed-up the
pipeline execution by finishing a branch of task computations before
starting new ones \TG{This is meant to reduce the memory footprint of the
application}. \TG{Here we should add a few comparison sentences between
Spark's and Dask's schedulers.} Dask has five main APIs: Dask Bags, Dask
Arrays, Dask Futures, Dask Delayed and DataFrames. We compared them all
except DataFrames as they are not adapted to the type of data we
considered. All APIs provide in-memory computing, data locality, and lazy
evaluation, however, none of them has fault tolerance \TG{Check fault
tolerance in Dask. Fine-grained lineage vs Spark's coarse grained lineage}.
We used Dask v1.1.4.

\href{https://docs.dask.org/en/latest/bag.html}{Dask Bags}
are parallel collections of Python objects, similar to Spark's RDD. Dask
Bags also offers optimized algorithms \TG{``optimized algorithms'' is
inaccurate: you should explain what they do} that mimic most of the
toolz~\cite{toolz} library. \href{https://docs.dask.org/en/latest/array.html}{Dask
Arrays} are used for the
processing of large arrays. They provide a parallel clone of the popular
NumPy library. \href{https://docs.dask.org/en/latest/delayed.html}{Dask
Delayed} supports arbitrary
tasks that does not fit in the Array, DataFrame or MapReduce \TG{Do you
mean Bags?} frameworks. Finally, \href{https://docs.dask.org/en/latest/futures.html}{Dask
Futures} is similar to Dask
Delayed as it supports arbitrary tasks, but it operates in real-time rather
than lazily.

@@

Mathieu Dugr\'e was funded by a Undergraduate Summer Research Assistant
award from the National Science and Engineering Research Council. We warmly
thank Compute Canada and \TG{Westgrid?} for providing the cloud
infrastructure used in these experiments, and the McGill Center for
Integrative Neurocience for giving us access to their cloud allocation.
